---
title: "Assignment 01"
output: html_document
date: "2025-05-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

By including this statement, we the authors of this work, verify that:

• We hold a copy of this assignment that we can produce if the original is lost or damaged.

• We hereby certify that no part of this assignment/product has been copied from any other student’s
work or from any other source except where due acknowledgement is made in the assignment.

• No part of this assignment/product has been written/produced for us by another person except where such collaboration has been authorised by the subject lecturer/tutor concerned.

• We are aware that this work may be reproduced and submitted to plagiarism detection software programs for the purpose of detecting possible plagiarism (which may retain a copy on its database for future plagiarism checking).

• We hereby certify that we have read and understand what the School of Computing, Engineering and Mathematics defines as minor and substantial breaches of misconduct as outlined in the learning
guide for this unit.


1. Write the code to analyse the review behaviour across user groups. The users should be grouped into 3 group: Veteran, Intermediate and New (based on their member since date) before 2017, between 2017-2022, and after 2022 respectively. Calculate the numbers of users, their average review stars and average number of reviews per user. Tabulate the data using kable or kableextra. Visualise the Average Review Stars by User Age Group. You are required to make sure you handle the NA value in your analysis. Explain your findings.

Solution:

```{r}
library(dplyr)
library(tidyverse)
library(kableExtra)
library(ggplot2)
data_users=read.csv("users.csv")
data_users <- data_users %>%
  mutate(
    users_group = case_when(
      data_users$member_since < as.Date("2017-01-01") ~ "Veteran",
      data_users$member_since >= as.Date("2017-01-01") & data_users$member_since <= as.Date("2022-12-31") ~ "Intermediate",
      data_users$member_since > as.Date("2022-12-31") ~ "New",
      TRUE ~ NA_character_
    )
  )
head(data_users)
user_stats <- data_users %>%
  summarise(
    # Total number of users
    total_users = n_distinct(data_users$user_id),
    
    # Average review stars per user
    avg_review_stars = mean(data_users$average_stars, na.rm = TRUE),
    
    # Average number of reviews per user
    avg_reviews_per_user = mean(data_users$review_count, na.rm = TRUE)
  )
colnames(user_stats)=c("Total Users","Average review stars per user","Average number of reviews per user")
```


```{r}
user_stats %>% kbl(caption = "Table for numbers of users and their reviews") %>%  kable_paper("hover",full_width=F,html_font = "Camberia")
```

```{r}
average_review_bygroup <- data_users %>% 
  group_by(users_group) %>%  
  summarise(
    avg_review_stars = mean(average_stars, na.rm = TRUE), 
    .groups = 'drop'
  )
# Clean and prepare data
plot_data <- average_review_bygroup %>%
  filter(
    !is.na(users_group),
    !is.na(avg_review_stars),
    is.numeric(avg_review_stars)
  ) %>%
  mutate(
    users_group = factor(users_group,
                        levels = c("New", "Intermediate", "Veteran"),)
  )

data_min <- min(plot_data$avg_review_stars)
data_max <- max(plot_data$avg_review_stars)
y_buffer <- (data_max - data_min) * 0.3  # 30% padding
barplot(plot_data$avg_review_stars,names.arg = plot_data$users_group,main = "Average Review Stars by User Group",
        ylab = "Average Stars",
       ylim = c(data_min - y_buffer, data_max + y_buffer))  # Focused y-axis)
text(x = 1:nrow(plot_data), 
     y = plot_data$avg_review_stars,
     labels = round(plot_data$avg_review_stars, 3),
     pos = 3, cex = 0.8)
```



2. Write the code to analyse the average reviews star by State. Calculate the average review star, the
number of reviews and the number of unique users. Visualise the Average Review Stars by State.
You are required to make sure you take care of the NA value in your analysis. Elaborate on the
findings.

Solution:


```{r}
data_bs=read.csv("businesses.csv")
head(data_bs)
```
```{r}
data_by_state=data_bs %>% filter(!is.na(state) & state != "") %>% 
  group_by(state) %>% 
  summarise(
  aver_review =mean(business.avg.stars, na.rm=TRUE)
)
head(data_by_state)
```

```{r}
business_stas <-data_bs  %>%
  summarise(
    # Total number of users
    total_business = n_distinct(data_bs$business_id),
    
    # Average review stars per user
    avg_review_stars = mean(data_bs$business.avg.stars, na.rm = TRUE),
    
    # Average number of reviews per user
    num_review = sum(data_bs$review_count, na.rm = TRUE)
  )
head(business_stas)
```


```{r}
data_by_state$state=as_factor(data_by_state$state)
ggplot(data_by_state, aes(x = state, y = aver_review,size = aver_review))+geom_point(alpha=0.5)+theme_minimal()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) # Rotate axis label
```









3. Write the code to analyse the top users and their behaviours. First, identify the top 10 users by the
review count. For those top 10 users, calculate their average review stars. Tabulate the summary of
the data (kable/kableextra). You are required to make sure you handle the NA value in your analysis.
Visualise their rating distrubtion using ggplot2 - boxplot. Discuss your findings.


Solution:

```{r}
top_10=data_users %>% arrange(desc(data_users$review_count)) %>% slice(1:10)
head(top_10)
top_10$average_review=top_10$review_count
```


4.Write the code to analyse if there is a major difference between the review behavior of users who
joined before and after 2020. For these 2 groups of users, compare their star rating behaviour and
the length of the reviews (number of charaters in the review text). You are required to make sure
you handle the NA value in your analysis. Visualise the average review length by the two groups.
Discuss your findings.


Solution:

```{r}
user_cat=data_users %>% mutate(
  member_since=as.Date(member_since),
  Before_2020=if_else(
      member_since < as.Date("2020-01-01"), 
      TRUE, 
      FALSE
    )
) %>% filter(!is.na(Before_2020))
head(user_cat)
```


```{r}
data_rv=read_csv("reviews.csv",show_col_types = FALSE)
head(data_rv)
join_table=data_rv %>% inner_join(user_cat,by="user_id") %>% mutate(review_length=nchar(text)) %>% filter(!is.na(stars),!is.na(review_length))

```


```{r}
review_by_group=join_table %>% group_by(Before_2020) %>% 
  summarise(
              average_stars_rating=mean(stars,na.rm=TRUE),
              avg_review_length = mean(review_length, na.rm = TRUE),    
              n_reviews = n(),
              n_unique_users = n_distinct(user_id))

head(review_by_group)
```


```{r}
colnames(review_by_group)=c("Join before 2020","Average star rating","Average length of review ", "Number of reviews","Number of users")
review_by_group %>% kbl(caption = "Summary by Group") %>%  kable_paper("hover",full_width=F,html_font = "Camberia")
```


```{r}
ggplot(join_table, aes(x = Before_2020, y = review_length)) +
  geom_boxplot(fill = "white", alpha = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  labs(title = "Review Length Distribution by User Join Period",
       x = "User Group",
       y = "Review Length (characters)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```





