---
title: "Assignment 01"
output: html_document
date: "2025-05-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

By including this statement, we the authors of this work, verify that:

• We hold a copy of this assignment that we can produce if the original is lost or damaged.

• We hereby certify that no part of this assignment/product has been copied from any other student’s
work or from any other source except where due acknowledgement is made in the assignment.

• No part of this assignment/product has been written/produced for us by another person except where such collaboration has been authorised by the subject lecturer/tutor concerned.

• We are aware that this work may be reproduced and submitted to plagiarism detection software programs for the purpose of detecting possible plagiarism (which may retain a copy on its database for future plagiarism checking).

• We hereby certify that we have read and understand what the School of Computing, Engineering and Mathematics defines as minor and substantial breaches of misconduct as outlined in the learning
guide for this unit.


```{r}
#load necessary libraries 
library(dplyr)
library(tidyverse)
library(kableExtra)
library(ggplot2)
#Load data
data_users=read.csv("users.csv")
data_bs=read.csv("businesses.csv")
data_rv=read_csv("reviews.csv",show_col_types = FALSE)
```

1. Write the code to analyse the review behaviour across user groups. The users should be grouped into 3 group: Veteran, Intermediate and New (based on their member since date) before 2017, between 2017-2022, and after 2022 respectively. Calculate the numbers of users, their average review stars and average number of reviews per user. Tabulate the data using kable or kableextra. Visualise the Average Review Stars by User Age Group. You are required to make sure you handle the NA value in your analysis. Explain your findings.

Solution:

```{r}
#Summary and the first 10 rows of data_users
summary(data_users)
head(data_users)
```

```{r}
#Create new field for Veteran, Intermediate, and New
data_users <- data_users %>%
  mutate(
    users_group = case_when(
      data_users$member_since < as.Date("2017-01-01") ~ "Veteran",
      data_users$member_since >= as.Date("2017-01-01") & data_users$member_since <= as.Date("2022-12-31") ~ "Intermediate",
      data_users$member_since > as.Date("2022-12-31") ~ "New",
      TRUE ~ NA_character_
    )
  )
```

```{r}
#Create new dataframe for calculated statistics 
user_stats <- data_users %>%
  summarise(
    # Total number of users in the first collumn
    total_users = n_distinct(data_users$user_id),
    
    # Average review stars per user in the second collumn
    avg_review_stars = mean(data_users$average_stars, na.rm = TRUE),
    
    # Average number of reviews per user in the third collumn
    avg_reviews_per_user = mean(data_users$review_count, na.rm = TRUE)
  )
#Change the name of collumns for easy-reading
colnames(user_stats)=c("Total Users","Average review stars per user","Average number of reviews per user")
```


```{r}
#Set the title for the formated graph with front of Camberia, expanding across paper, hover using to change color in html/ webpage
user_stats %>% kbl(caption = "Table for numbers of users and their reviews") %>%  kable_classic("hover",full_width=F,html_font = "Camberia")
```

```{r}
#group avg_review_stars by groups of user
average_review_bygroup <- data_users %>% 
  group_by(users_group) %>%  
  summarise(
    avg_review_stars = mean(average_stars, na.rm = TRUE), 
    .groups = 'drop'
  )
# Clean and prepare data
plot_data <- average_review_bygroup %>%
  filter( #filter out/ collect TRUE value from statement 
    !is.na(users_group), #filter out NA in users_group
    !is.na(avg_review_stars),#filter out NA in avg_review_stars
    is.numeric(avg_review_stars) #checking the structure of the avg_review_stars as numeric
  ) %>%
  mutate( #double-checking/add new group of users_group as factor
    users_group = factor(users_group,
                        levels = c("New", "Intermediate", "Veteran"),)
  )
ggplot(plot_data, aes(x = users_group, y = avg_review_stars)) +
   geom_col(fill = "black") +
  geom_text(aes(label = round(avg_review_stars, 2)), vjust = -0.5)+
  labs(
    title = "Average Review Stars by User Group",
    x = "User Group",
    y = "Average Review Stars"
  ) +
  theme_minimal()
```



2. Write the code to analyse the average reviews star by State. Calculate the average review star, the
number of reviews and the number of unique users. Visualise the Average Review Stars by State.
You are required to make sure you take care of the NA value in your analysis. Elaborate on the
findings.

Solution:


```{r}
#First 10 rows of data_bs
head(data_bs)
```
```{r}
#Create new dataframe for review each state
data_by_state=data_bs %>% #Filter out NA state and ungiven state
  filter(!is.na(state) & state != "") %>% 
  group_by(state) %>% #Group by state
  summarise(#Return aver_review as a new collumn for mean for each state
  aver_review =mean(business.avg.stars, na.rm=TRUE)
)
#First 10 rows of data_by_state
head(data_by_state)
```

```{r}
# Create new dataframe for summary statistics
business_stats <- data_bs %>%
  summarise(
    # Total number of businesses
    total_business = n_distinct(business_id),
    
    # Average review stars per business
    avg_review_stars = mean(business.avg.stars, na.rm = TRUE),
    
    # Number of reviews of all businesses
    num_reviews = sum(review_count, na.rm = TRUE)
  )

# Add user count from a separate data_users dataframe
user_count <- data_users %>%
  summarise(user_count = n_distinct(user_id))

# Combine the user count with business stats
business_stats <- business_stats %>%
  mutate(user_count = user_count$user_count) %>%
  select(total_business, user_count, avg_review_stars, num_reviews)

# Rename columns
colnames(business_stats) <- c(
  "Total number of businesses",
  "Total number of users",
  "Average review stars per business",
  "Total number of reviews"
)

# Create formatted table
business_stats %>%
  kbl(caption = "Table: Summary of Businesses, Users, Review Stars, and Total Reviews") %>%
  kable_classic("hover", full_width = F, html_font = "Cambria")
```


```{r}
data_by_state$state=as_factor(data_by_state$state)
ggplot(data_by_state, aes(x = state, y = aver_review,size = aver_review))+geom_point(alpha=0.5)+theme_minimal()+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) # Rotate axis label
```









3. Write the code to analyse the top users and their behaviours. First, identify the top 10 users by the review count. For those top 10 users, calculate their average review stars. Tabulate the summary of
the data (kable/kableextra). You are required to make sure you handle the NA value in your analysis.
Visualise their rating distrubtion using ggplot2 - boxplot. Discuss your findings.


Solution:

```{r}
top_10=data_users %>% arrange(desc(data_users$review_count)) %>% slice(1:10)
head(top_10)
top_10$average_review=top_10$review_count
```


4.Write the code to analyse if there is a major difference between the review behavior of users who
joined before and after 2020. For these 2 groups of users, compare their star rating behaviour and
the length of the reviews (number of charaters in the review text). You are required to make sure
you handle the NA value in your analysis. Visualise the average review length by the two groups.
Discuss your findings.


Solution:

```{r}
user_cat=data_users %>% mutate(
  member_since=as.Date(member_since),
  Before_2020=if_else(
      member_since < as.Date("2020-01-01"), 
      TRUE, 
      FALSE
    )
) %>% filter(!is.na(Before_2020))
head(user_cat)
```


```{r}
head(data_rv)
join_table=data_rv %>% inner_join(user_cat,by="user_id") %>% mutate(review_length=nchar(text)) %>% filter(!is.na(stars),!is.na(review_length))

```


```{r}
review_by_group=join_table %>% group_by(Before_2020) %>% 
  summarise(
              average_stars_rating=mean(stars,na.rm=TRUE),
              avg_review_length = mean(review_length, na.rm = TRUE),    
              n_reviews = n(),
              n_unique_users = n_distinct(user_id))

head(review_by_group)
```


```{r}
colnames(review_by_group)=c("Join before 2020","Average star rating","Average length of review ", "Number of reviews","Number of users")
review_by_group %>% kbl(caption = "Summary by Group") %>%  kable_paper("hover",full_width=F,html_font = "Camberia")
```


```{r}
ggplot(join_table, aes(x = Before_2020, y = review_length)) +
  geom_boxplot(fill = "white", alpha = 0.7) +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3, color = "red") +
  labs(title = "Review Length Distribution by User Join Period",
       x = "User Group",
       y = "Review Length (characters)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```





